---
title: "test_doc"
output: html_document
---
### Loading required libraries
First, we'll load all required libraries

```{r libraries, results='hide', message=FALSE, warning=FALSE}
library(HGNChelper)
library(openxlsx)
library(expss)
library(ggpubr)
library(dplyr)
library(scales)

```

```{r universeskip,include=FALSE}
load("output/Data/universe_df.rda")
```

### Initialising data frame of all protein-coding genes

Next, we'll initialise our data frame of all protein-coding genes, universe_df, which we'll be using throughout the project. 

To do this, we'll load in HGNC's spreadsheet of protein-coding genes, downloaded from [their website](https://www.genenames.org/cgi-bin/statistics), updating gene symbols using, and taking only the columns we want. 

```{r universe_init, eval = FALSE}
#read protein-coding gene list
hgnc<- read.csv("Gene_lists/Universe/gene_with_protein_product.txt",sep = "\t", comment.char = "#",stringsAsFactors = FALSE)

#take only needed columns
hgnc$symbol<- checkGeneSymbols(hgnc$symbol)[[3]]
universe_df <- data.frame(hgnc$hgnc_id,hgnc$symbol,hgnc$name,hgnc$gene_family,hgnc$location,hgnc$mgd_id)
names(universe_df) <- c("hgnc_id","gene","gene_name","gene_family","location","mgd_id")
rm(hgnc)

```

